### 2024-04-01 manuscript.Rmd
structuring question and some discussion

In this study, we investigate the interspecific variation in the number of vertebrae to test the hypothesis that changes in developmental patterns underpin the evolution of different body shapes in Australian typhlopid snakes.

Specifically, we test two predictions:
(1) narrow-and-long species develop from postnatal ontogenetic changes where vertebral numbers are not significantly different between narrow-and-long species compared to wide-and-robust species; but are elongated. 
(2) Narrow-and-long blindsnakes develop via changes in the embryonic and postnatal stages where narrow-and-long species have greater number of vertebrae than wide-and-robust species.

To test these predictions, we quantified the number of pre- and post-clocal vertebra across 37 species of Australian blindsnakes.

@@

We extracted 37 lineages of Australian blindsnakes from the time-calibrated phylogeny of Australian blindsnakes and outgroups presented in [@tiatragulAustralia2023]. The subset phylogeny represents 35/47 formally described species of _Anilios_ along with two additional genetically distinct lineages: one eastern population of "_A. grypus_" and one eastern population of "_A. ligatus_" [@tiatragulAustralia2023]. This phylogeny was inferred using the shortcut coalescent method in ASTRAL III [@zhangASTRALIIIPolynomialTime2018] based on 4930 maximum likelihood gene trees estimated using IQTREE [@minhIQTREE2020]. Divergence dating was estimated using MCMCTree from the PAML package [@yangPAML2007] based on 27 squamate reptile fossil calibrations [@tiatragulAustralia2023].

@@

@@

We found a weakly positive but non statistically significant relationship between maximum number of total vertebrae and maximum SVL (PGLS, slope=0.28, y=256, R^2^=0.02, F=0.72, Z=0.29, P=0.42), indicating an absence of pleomerism in Australian blindsnakes (Fig. \@ref(fig: fig-tbl-vert)).

However, we found that vertebrae ratio is postively correlated with aspect ratio (PGLS, slope=-0.0048, y=0.73, R^2^=0.17, F=7.09,, Z=2.2, P=0.011), indicating that species that are more narrow-and-wide usually have greater total vertebrae to total length ratios than species that are wide-and-robust (Fig. \@ref(fig:fig-vert-width-ratios)).

@@

@@

@@

## Miscellaneous 
Traditionally, ventral scale counts have been used as a proxy for vertebral number as they correlate for most alethinophidian snakes [@alexanderPattern1966]. We found that this is (not) the case in blindsnakes.

>>>>

In this study, we investigate the prevalence of pleomerism among Australian blindsnakes and test hypotheses that deviations from pleomerism may be explained by different body shapes or other ecological factors among Australian typhlopid snakes. The specific questions we addressed are:

1. How prevalent is pleomerism among blindsnakes?

2. Can deviations from pleomerism be explained by body shape in a genus of Australian blindsnakes?

3. Do vertebral number and body size vary with mean annual temperature?

<!-- Specifically, we test two predictions: -->
<!-- (1) narrow-and-long species develop from postnatal ontogenetic changes where vertebral numbers are not significantly different between narrow-and-long species compared to wide-and-robust species.  -->
<!-- (2) Narrow-and-long blindsnakes develop via changes in the embryonic and postnatal stages where narrow-and-long species have greater number of vertebrae than wide-and-robust species.  -->

To answer these questions we gathered data on the number of vertebrae across XX species of Australian blindsnakes.

@@

We extracted 37 lineages of Australian blindsnakes from the time-calibrated phylogeny of Australian blindsnakes and outgroups presented in [@tiatragulAustralia2023]. The subset phylogeny represents 35/47 formally described species of _Anilios_ along with two additional genetically distinct lineages: one eastern population of "_A. grypus_" and one eastern population of "_A. ligatus_" [@tiatragulAustralia2023]. This phylogeny was inferred using the shortcut coalescent method in ASTRAL III [@zhangASTRALIIIPolynomialTime2018] based on 4930 maximum likelihood gene trees estimated using IQTREE [@minhIQTREE2020]. Divergence dating was estimated using MCMCTree from the PAML package [@yangPAML2007] based on 27 squamate reptile fossil calibrations [@tiatragulAustralia2023]. This phylogeny is was used for all phylogenetic comparative analyses.

@@

### Prevalence of pleomerism among blindsnakes

@@

@@

For the Australian blindsnakes, we found a weakly positive but non statistically significant relationship between maximum number of total vertebrae and maximum SVL (PGLS, slope=0.28, y=256, R^2^=0.02, F=0.72, Z=0.29, P=0.42), indicating an absence of pleomerism in Australian blindsnakes (Fig. \@ref(fig: fig-tbl-vert)). This suggest that maximum length is acheived by elongating each vertebrae rather than by adding additional vertebrae

### Relationship between pleomerism and body shape
However, we found that vertebrae ratio is postively correlated with aspect ratio (PGLS, slope=-0.0048, y=0.73, R^2^=0.17, F=7.09,, Z=2.2, P=0.011), indicating that species that are more narrow-and-wide usually have greater total vertebrae to total length ratios than species that are wide-and-robust (Fig. \@ref(fig:fig-vert-width-ratios)). This is consistent with colubrids and viperids, where species that are more elongated tend to have greater number of vertebrae than stouter ones [@lindellEvolution1994].

@@

### Correlation with mean annual temperature among Australian blindsnakes

@@

Summary about what the paper found

Not so surprising but shows that pleomerism does not explain variation in body size in this continental radiation of blindsnakes. Only real giant taxa like *Rhinotyphlops* and *Afrotyphlops* tend to break the rule. However, when fitted in our PGLS, we found that even they are not exceptional???

Body shape varied with number of vertebrae in similar ways as previously reported in vipers and colubrids [@lindellEvolution1994]. We can speculate that species that are narrow-and-long may be using pre-existing burrows rather than digging their own burrow. Narrow-and-wide phenotypes are common among Leptotyphlopidae, which are thought to use existing burrows rather than dig their own burrow [@herrelBurrowing2021]. The Australian group is quite exceptional because within the same genus, their is great variation in aspect ratio so species that evolved more narrow-and-long bodies may be converging on the lifestyle with other blindsnakes that have this narrow-and-long phenotype. On the other hand, large robust-and-wide species are expected to be able to generate much more force and make their own burrows. A previous study (Tiatragul et al. unpublished) found that species that robust-and-wide species tend to occupy areas in Australia that have easier to penetrate soil compared to narrow-and-long species that tend to occupy areas with more difficult to penetrate soil. However, more detailed microhabitat observations is required to determine if body shape is related to the environment in which they burrow.

Temperature seem to have XXX effect. Common garden experiments for different species of snakes can tell us if incubation temperature and thus rate of development influences number of vertebrae. We would expect higher temperature to accelerate development (CITE anolis papers, turtles, etc), so less time for embryonic development.

## Miscellaneous 
Traditionally, ventral scale counts have been used as a proxy for vertebral number as they correlate for most alethinophidian snakes [@alexanderPattern1966]. We found that this is (not) the case in blindsnakes. Blindsnakes are generally small enough to fit in an X-ray scanner so more people should do it.
